Unsupervised training of maximum-entropy models for lexical selection in rule-based machine translation

### 2015-04-12 2015-eamt-melsa/2015-eamt-melsa.tex
Changes made or indicated
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59722 72bbbca6-d526-0410-a7d9-f06f51895060

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{multirow}
\usepackage{url}
\usepackage{natbib}
\usepackage{color}
\usepackage[small,bf]{caption}
\usepackage{microtype} % a.k.a. conference paper magic

@@

\author{
\phantom{Francis M.\ Tyers} \\
\phantom{HSL-fakultehta,} \\
\phantom{UiT Norgga \'{a}rktala\v{s} universitehta,} \\
\phantom{N-9018 Romsa} \\
 \And
\phantom{Felipe S\'anchez-Mart\'inez} \\
\phantom{Dept.\ Lleng.\ i Sist.\ Inform.,} \\
\phantom{Universitat d'Alacant,}\\
\phantom{E-03071 Alacant}\\
 \And
\phantom{Mikel L.\ Forcada} \\
\phantom{Dept.\ Lleng.\ i Sist.\ Inform.,} \\
\phantom{Universitat d'Alacant,}\\
\phantom{E-03071 Alacant}\\
}

@@

\section{Introduction}

Corpus-based MT has been the primary research direction in the field
of machine translation in recent years. However, rule-based MT (RBMT)
systems are still being developed, and there are many successful
commercial and non-commercial systems. One reason for the continued
development of RBMT systems is that in order to be successful,
corpus-based MT requires parallel corpora in the order of tens of
millions of words. Although for some language pairs these exist, they
only exist for a fraction of the world's languages.

@@

\subsection{Prior work}
%@@@ \comment{This section should be ironed out a bit.}  
%In their pioneering work, 
\cite{dagan1994word} used the term \emph{word sense disambiguation} to
refer to what is actually target word selection in machine
translation; they used a parser to identify some syntactic relations
such as subject--object or subject--verb.  After generating all the
possible translations for a given input sentence using an ambiguous
bilingual dictionary, they extract the syntactic tuples from the
target language and count the frequency in a previously-trained
target-language model of tuples.  They use maximum-likelihood
estimation to calculate the probability that a given target language
tuple is the translation of a given source language tuple, with an
automatically determined confidence threshold.

@@

\section{Method}

%\todo[inline]{FSM: In the experiments a development corpus is
%  mentioned. For what is it used? This should be explain in this
%  section.}

@@

Let the probability of a word $t$ being the translation of a
word $s$ in an SL context $c$ be $p_s(t|c)$. In principle, this value
could be estimated directly from the available corpora for every
combination of $(s, t, c)$.  This would however present two questions:
(1) how should the relevant contexts be chosen? and (2) what should be
done when $(s, t, c)$ is not found in the corpus?
A maximum-entropy model answers both of these questions. It allows the
contexts that we consider to be linguistically interesting to be
defined \emph{a priori} and then integrate these seamlessly into a
probabilistic model \citep{Manning99b}.  In answer to the second
question, a maximum-entropy model assumes nothing about what is not
present in the training data. That is, if there is no information in
the training data, then it assumes that all outcomes ---that is, all
possible translations--- are equally likely.  As previously mentioned,
the principle of maximum entropy has been applied to the problem of
lexical selection before; in particular, \cite{berger1996} cast the
problem of lexical selection in statistical MT as a classification
problem. They learn a separate maximum-entropy classifier for each SL
word form,
%\comment{[:we do too, integrate this later]} 
using SL context to
distinguish between possible translations. These classifiers are then
incorporated into the translation model of their word-based SMT
system.
%\comment{[how is our work different?]}  
In their approach, a
classifier consists of a set of binary feature functions and
corresponding weights for each feature. 
In both \cite{berger1996} and our method, features are defined in the
form $h_k^{s}(t, c)$, where $t$
is a translation, and $c$ is an SL context. One difference is that \cite{berger1996} take $s$, $t$
and $c$ to be based on word forms, whereas in our method they are based
on lemma forms.
A feature where the Spanish word \emph{pez} (`fish' as a living animal) is seen as the
translation of \emph{arrain} (`fish') in the context \emph{arrain
  handi} `big fish' would therefore be defined as:
%\comment{FSM: I have introduced the index $k$
%  because several features are defined for a SL word $s$.} 
\begin{equation}
h_k^{\mathrm{arrain}}(t, c) = \left \{ \begin{matrix}
                 ~~1 & \mbox{if } 
\left\{ \begin{array}{c}
t~ =~ \textit{pez}~ \\ \mathrm{and}~ \\ \textit{handi}~ \mathrm{follows}~ \textit{arrain} \end{array} \right. \\
                 ~~0 & \mbox{otherwise} \end{matrix} \right.
\label{eq:feat-func}
\end{equation}
This feature considers a context of zero words to the left of the
problem word and one word to the right of it.

@@

As a result of training, each feature in the classifier is assigned a
weight $\lambda_k^s$. Combining these weights of active features
as in equation~(\ref{eq:max-ent-sum}) yields the probability of a
translation $t$ for word $s$ in context $c$.

\begin{equation}
p_s(t|c) = \frac{1}{Z^s(c)} \exp{ \sum_{k=1}^{n_F} \lambda_k^s h_k^s(t, c)}
\label{eq:max-ent-sum}
\end{equation}

@@

The input to our method consists of a collection of samples,
$\mathcal{G} = (S, G)$, where $S = (s_1, s_2, \ldots s_{|S|})$ is a
sequence of SL words, and $G = \{g_1, g_2, \ldots g_{|G|}\}$ is a set
of possible \emph{lexical-selection paths}. A lexical-selection path
$g = (t_1, t_2, \ldots, t_{|S|})$ is a sequence of lexical-selection
choices of those SL words, where \(t_i \in
T_s(s_i)\). % and \(T_s(\cdot)\)
%is the set of all
%translations for a source word. 
This is produced in the first stages of RBMT, just after morphological
analysis, part-of-speech tagging, and bilingual dictionary lookup, and
before any structural transfer takes place (we will call this
\emph{pre-lexsel}). In our model, it is after these first stages that
lexical selection (\emph{lexsel}) occurs. After lexical selection,
structural transfer and generation take place; a function
$\tau(g_i,S)$ represents the result of these last stages, which we
will call \emph{post-lexsel}, and returns a finished translation of a
specific lexical-selection path $g_i$ of sentence
$S$. Figure~\ref{fig:lexselschema} shows this process
schematically. %%%\todo{Can we try to bring the figure closer?}

%Taking our previous example for the parallel corpus, Table~\ref{table:ex-monocorp} 
%presents an example of five possible samples.

@@

suitable normalisation:

@@

\section{Experimental setting}

@@

Evaluation will be performed using four language pairs. These pairs
have been selected as they include languages with different
morphological complexity, and different amounts of resources available
--- although for all pairs there is a parallel corpus available
for evaluation (see Section \ref{ss:measures}). 
% Table~\ref{table:lang-pair-stats} 
%presents statistics about the development status of each of the language 
%pairs chosen, including the number of bilingual dictionary entries, coverage of
%test corpora and translational ambiguity of the bilingual dictionary. Translation
%ambiguity is provided for both all words, and for only ambiguous open-category 
%words. In the experiments, lexical-selection rules are learnt for the open categories
%of adjective, noun and verb.

@@

\item[Breton--French:] Development of the Breton--French pair has been
  described in \citep{tyers10b}. The bilingual % tyers09b
  dictionaries were not built with polysemy in mind from the outset,
  but some entries were added later to start work on lexical
  selection. The version used in this paper is SVN revision
  41375.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-br-fr}}
%  This is the only available machine translation system between Breton
%  and French. It has been developed part-time over a number of years.

@@

\item[English--Spanish:] The English--Spanish pair was developed from
  a combination of the English--Catalan and Spanish--Catalan pairs
  over a period of around four months. The pair
%has been used in other
%  lexical selection experiments \citep{sanchez07a}\comment{FSM: Sure?
%    It was on en-ca and in that paper no experimental results are
%    shown. This paper would be better mentioned in the introduction
%    (it mentions the use of TL info for lexical selection.},
  contains a number of entries in the bilingual dictionary with more
  than one translation. %The most-frequent translation, as judged by
%  the language pair developer, is marked as the \emph{default}
%  translation and non-default translations were added according to
%  frequency. \todo{Default is explained for some LPs only, why?}  
  The
  version used in this paper is SVN revision
  41387.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-en-es}}
\end{description}

@@

% see section~\ref{sec:ler}. 
\begin{figure*}
\begin{center}
\setlength\tabcolsep{5.2pt}%
 \begin{tabular}{|l|llllll|}
   \hline
   \multicolumn{7}{|c|}{L'estiu \'{e}s una estaci\'{o} llarga}\\
   \hline
   $S$        & \emph{el} & \emph{estiu} & \emph{ser} & \emph{un} & \emph{estaci\'{o}} & \emph{llarg} \\
   \hline
   $T_s(s_i)$        & \{the\} & \{summer\} & \{be\} & \{a\} & \{station, season\} &  \{long, lengthy\} \\
   \hline 
   $T_r(s_i)$        & \{the\} & \{summer\} & \{be\} & \{a\} &  \{season\} & \{long\} \\
   \hline
   $T_t(s_i)$        & \{the\} & \{summer\} & \{be\} & \{a\} &  \{station\} & \{long\} \\
   \hline
   $\textrm{amb}(s_i)$ & 0     & 0          & 0      & 0     & 1            & 1 \\
   \hline
   $\mathrm{diff}(T_r(s_i), T_t(s_i))$ & 0     & 0          & 0      & 0     & 1            & 0 \\
   \hline
 \end{tabular}
 \caption{An example input sentence in Catalan and the three sets of
   English translations used for calculating the lexical-selection
   error rate. The source sentence $S = (s_1, s_2, \ldots, s_{|S|})$
   has two ambiguous words, \emph{estaci\'{o}} and \emph{llarg}. There is one difference
   between the reference set $T_r(s_i)$ and the test set $T_t(s_i)$ of
   translations; thus, the error rate for this sentence is
   50\%.}
\label{fig:ler-example}
\end{center}
\end{figure*}

@@

The lexical-selection error rate is the fraction of times the given
system chooses a translation for a word which is not  the one found in an
annotated reference. The process uses a source-language sentence, $S =
(s_1, s_2, \ldots, s_{|S|})$ and three functions. The first function,
$T_s(s_i)$, returns all possible translations of $s_i$ according to
the bilingual dictionary. The second function, $T_t(s_i)$, returns the
translations of $s_i$ selected by the lexical-selection module:
$T_t(s_i) \subseteq T_s(s_i)$; and usually $|T_t(s_i)| = 1$. If the
lexical-selection module returns more than one translation, the first
translation is selected. 
%! \comment{FSM: How many times does this
%!  happens?. I would move this last sentence to an appendix.}  
The
function $T_r(s_i)$ returns the set of reference translations which
are acceptable for $s_i$ in sentence $S$.\footnote{Depending on how
  the reference is built, the set returned by $T_r(s_i)$ % may be
  % incomplete in the sense that it 
  may not include all possible
  acceptable translations.} For a
single sentence, we define the lexical selection error rate (LER) of
that sentence as
\begin{equation}
\mathrm{LER} = \frac{\sum_{i=1}^{|S|} \mathrm{amb}(s_i)\; \mathrm{diff}(T_r(s_i), T_t(s_i))}{\sum_{i=1}^{|S|} \mathrm{amb}(s_i)},
\label{eq:ler}
\end{equation}
where
\begin{equation}
\mathrm{amb}(s_i) = \left \{ \begin{matrix}
                 1 & \mbox{if } |T_s(s_i)| > 1   \\
                 0 & \mbox{otherwise} \end{matrix} \right.
\label{eq:amb}
\end{equation}
 tests if a word is ambiguous, and the  function 
\begin{equation}
\mathrm{diff}(T_r(s_i), T_t(s_i)) = \left \{ \begin{matrix}
                 1 & \mbox{if } T_r(s_i) \cap T_t(s_i) = \emptyset   \\
                 0 & \mbox{otherwise} \end{matrix} \right.
\label{eq:diff}
\end{equation}
states that there is a difference if the intersection between the set
of reference translations $T_r(s_i)$ and the set of translations from
the lexical selection module $T_t(s_i)$ is empty.  Recall that,
although $T_t(s_i)$ returns a set, this set will be a singleton, as
when the lexical-selection module returns more than one translation
the first one will be selected.\footnote{In practice this does not happen
as each ambiguous word has a \emph{default} translation.}

@@

The table in Figure~\ref{fig:ler-example} gives an
overview of the inputs. In the description it is assumed that the
reference translation has been annotated by hand. However, hand
annotation is a time-consuming process, and was not possible. A
description of how the reference was built is given in
section~\ref{sec:eval-corpus}.

\subsubsection{Machine translation performance} 
%! \comment{This part of the text needs compressing and merging} 
This is an extrinsic evaluation, which ideally would test how much the
system improves as regards an approximate measurement of final
translation quality in a real system. For this task, we use the
widely-used {\sc bleu} metric (see
Section~\ref{sec:eval-metrics}). This is not ideal for evaluating the
task of a lexical selection module as the performance of the module
will depend greatly on (a) the coverage of the bilingual dictionaries
of the RBMT system in question, and (b) the number of reference
translations. It is also worth noting that successful lexical
selections may not lead to successful translations due to inadequate
transfer of morphological features.
% For example, if
% we translate the Spanish phrase \emph{La gente le dice que no venga}
% `The people tell him not to come', where \emph{decir} can be
% translated as `say' or `tell' in English, and our machine translation
% system generates `The people \emph{tells} him not to come', it would
% be counted the same as if it were `The people \emph{says} him not to
% come', although the lexical selection made was more
% adequate. Generation errors --- where a word is found in the bilingual
% dictionary, but not in the target language morphological dictionary
% --- may also lead to the same problem.  
The {\sc bleu} metric \citep{papineni02} is
included only as it is commonly used to evaluate MT systems.

@@

  After this process, we selected 1,000 sentence pairs at random for
  testing (\texttt{test}), 1,000 for development
  (\texttt{dev})\footnote{The development corpus was used for checking
    the value for frequency pruning of features.} and left the
  remainder for training.
%This is a smaller
%number of sentences than used in typical evaluations (such as the WMT
%series, \cite{callisonburch2012}), but was motivated by the fact that
%the smallest corpus (for Breton--French) after sentence discarding was
%only little over 4,000 sentences long. 
  Table~\ref{table:input-corp} gives statistics about the size of the
  input corpora, and how many sentences were left after processing for
  testing, training and development. \todo{Why do ambiguity numbers
    change so much from the paper?} Table~\ref{table:test-corp} gives
  information about the test corpora. 
%! \todo{FSM: One may wonder which
%!    would be the results when the TL side of the training corpus is
%!    used to train the models in the usual way.}

@@

As we are working with binary features, we use the implementation of
generalised iterative scaling available in the {\sc
  yasmet}\footnote{\url{http://www-i6.informatik.rwth-aachen.de/web/Software/YASMET.html};
  the compilable version we used is available as part of the Apertium
  \texttt{lex-tools} package,
  \url{http://downloads.sourceforge.net/project/apertium/apertium-lex-tools/apertium-lex-tools-0.1.0.tar.gz}.}
to calculate the feature weights.  After learning the feature sets and
weights, we compute the evaluation measures described in
section~\ref{ss:measures}. There is an option to remove events 
$(s, t, c)$ which occur
less than a certain number of times in the training corpus. This is
referred to as the feature pruning frequency threshold --- features
occuring less than the threshold are discarded. The value was set
experimentally. Values of between two and seven were tested, and the
ones which provided the best improvement on the development corpus
were selected; they happen to come close to the rule-of-thumb value of
five that \citet[p.~596]{Manning99b} found to be effective.
Table~\ref{table:maxent-features} shows the number of features that
have eventually been used for each language pair.

@@

\section*{Acknowledgements}

\emph{Removed for review}

>>>>

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{multirow}
\usepackage{url}
\usepackage{natbib}
\usepackage{color}
\usepackage[small,bf]{caption}
\usepackage[backgroundcolor=lightgray]{todonotes}
\newcommand{\mytodo}[1]{\todo[inline]{#1}}
\newcommand{\todomlf}[1]{\renewcommand{\baselinestretch}{0.55}\todo{\scriptsize\sf MLF: #1}\renewcommand{\baselinestretch}{1.0}}

\usepackage{microtype} % a.k.a. conference paper magic

@@

\author{
Francis M.\ Tyers \\
HSL-fakultehta, \\
UiT Norgga \'{a}rktala\v{s} universitehta, \\
N-9018 Romsa \\
 \And
Felipe S\'anchez-Mart\'inez \\
Dept.\ Lleng.\ i Sist.\ Inform., \\
Universitat d'Alacant,\\
E-03071 Alacant\\
 \And
Mikel L.\ Forcada \\
Dept.\ Lleng.\ i Sist.\ Inform., \\
Universitat d'Alacant,\\
E-03071 Alacant\\
}

@@

\section{Introduction}
\todomlf{R1 has a general comment: ``watch out for long sentences''}
Corpus-based MT has been the primary research direction in the field
of machine translation in recent years. However, rule-based MT (RBMT)
systems are still being developed, and there are many successful
commercial and non-commercial systems. One reason for the continued
development of RBMT systems is that in order to be successful,
corpus-based MT requires parallel corpora in the order of tens of
millions of words. Although for some language pairs these exist, they
only exist for a fraction of the world's languages.

@@

\subsection{Prior work}
%@@@ \comment{This section should be ironed out a bit.}  
%In their pioneering work, 
\cite{dagan1994word} used the term \emph{word sense disambiguation} to
refer to what is actually target word selection in machine
translation; they used a parser to identify syntactic relations
such as subject--object or subject--verb.  After generating all the
possible translations for a given input sentence using an ambiguous
bilingual dictionary, they extract the syntactic tuples from the
target language and count the frequency in a previously-trained
target-language model of tuples.  They use maximum-likelihood
estimation to calculate the probability that a given target language
tuple is the translation of a given source language tuple, with an
automatically determined confidence threshold.

@@

\section{Method}
\todomlf{R3 says: ``The technique clearly resembles the "Deciphering Foreign Language," (S. Ravi and K. Knight) or “EM Decipherment for Large Vocabularies “
(M. Nuhn, and H. Ney. EM) approaches, although these techniques were not meant for learning the disambiguation but for learning the translation dictionaries.
The proposed method only executes one EM iteration in contrast to these works. It would be interesting to highlight differences and resemblances with these deciphering approaches.''. Check and cite? Also, single iteration WTF?  }
%\todo[inline]{FSM: In the experiments a development corpus is
%  mentioned. For what is it used? This should be explain in this
%  section.}

@@

Let the probability of a word $t$ being the translation of a
word $s$ in a source-language context $c$ be $p_s(t|c)$. In principle, this value
could be estimated directly from the available corpora for every
combination of $(s, t, c)$.  This would however present two questions:
(1) how should the relevant contexts be chosen? and (2) what should be
done when $(s, t, c)$ is not found in the corpus?
A maximum-entropy model answers both of these questions. It allows the
contexts that we consider to be linguistically interesting to be
defined \emph{a priori} and then integrate these seamlessly into a
probabilistic model \citep{Manning99b}.  In answer to the second
question, a maximum-entropy model assumes nothing about what is not
present in the training data. That is, if there is no information in
the training data, then it assumes that all outcomes ---that is, all
possible translations--- are equally likely.  As previously mentioned,
the principle of maximum entropy has been applied to the problem of
lexical selection before; in particular, \cite{berger1996} cast the
problem of lexical selection in statistical MT as a classification
problem. They learn a separate maximum-entropy classifier for each SL
word form,
%\comment{[:we do too, integrate this later]} 
using SL context to
distinguish between possible translations. These classifiers are then
incorporated into the translation model of their word-based SMT
system.
%\comment{[how is our work different?]}  
In their approach, a
classifier consists of a set of binary feature functions and
corresponding weights for each feature. 
In both \cite{berger1996} and our method, features are defined in the
form $h_k^{s}(t, c)$, where $t$
is a translation, and $c$ is a source-language (SL) context.\todomlf{Defined SL which had not been defined} One difference is that \cite{berger1996} take $s$, $t$
and $c$ to be based on word forms, whereas in our method they are based
on lemma forms.
An example would be the following feature where the Spanish word \emph{pez} (`fish' as a living animal) is seen as the
translation of \emph{arrain} (`fish') in the context \emph{arrain
  handi} `big fish' would therefore be defined as:
%\comment{FSM: I have introduced the index $k$
%  because several features are defined for a SL word $s$.} 
\begin{equation}
h_\mathrm{+handi}^{\mathrm{arrain}}(t, c) = \left \{ \begin{matrix}
                 ~~1 & \mbox{if } 
\left\{ \begin{array}{c}
t~ =~ \textit{pez}~ \\ \mathrm{and}~ \\ \textit{handi}~ \mathrm{follows}~ \textit{arrain} \end{array} \right. \\
                 ~~0 & \mbox{otherwise} \end{matrix} \right.
\label{eq:feat-func}
\end{equation}
This feature considers a context of zero words to the left of the
problem word and one word (+ \emph{handi}) to the right of it.\todomlf{Removed \(k\) and used a more informative subscript}

@@

As a result of training, each of the \(n_F\) features \(h^s_k(t,c)\) in the classifier is assigned a
weight $\lambda_k^s$. \todomlf{Reworded to define \(n_F\)} Combining these weights of active features
as in equation~(\ref{eq:max-ent-sum}) yields the probability of a
translation $t$ for word $s$ in context $c$.

\begin{equation}
p_s(t|c) = \frac{1}{Z^s(c)} \exp{ \sum_{k=1}^{n_F} \lambda_k^s h_k^s(t, c)}
\label{eq:max-ent-sum}
\end{equation}
\todomlf{R3 complains: `` Probability notation is confusing. I would suggest the use of \(p(t | c,s)\) instead of \(p_s(t|c)\), that will simplify notation. For instance instead of \(Z^s (c)\) we would have \(Z(c,s)\).''. Ignore or take onboard?}

@@

The input to our method consists of a collection of samples,
$\mathcal{G} = (S, G)$, where $S = (s_1, s_2, \ldots s_{|S|})$ is a
sequence of SL words, and $G = \{g_1, g_2, \ldots g_{|G|}\}$ is a set
of possible \emph{lexical-selection paths}. A lexical-selection path
$g = (t_1, t_2, \ldots, t_{|S|})$ is a sequence of lexical-selection
choices of those SL words, where \(t_i\) is an element of \(T_s(s_i)\), the set of possible translations of \(s_i\).\todomlf{R1 asked for this, even if redundant, and says: ``can one source word be translated as multiple words? What about multi-word expressions that cannot be translated word-by-word or source words that do not have a TL equivalent'', etc.} % and \(T_s(\cdot)\)
%is the set of all
%translations for a source word.
This is produced in the first stages of RBMT, just after morphological
analysis, part-of-speech tagging, and bilingual dictionary lookup, and
before any structural transfer takes place (we will call this
\emph{pre-lexsel}). In our model, it is after these first stages that
lexical selection (\emph{lexsel}) occurs. After lexical selection,
structural transfer and generation take place; a function
$\tau(g_i,S)$ represents the result of these last stages, which we
will call \emph{post-lexsel}, and returns a finished translation of a
specific lexical-selection path $g_i$ of sentence
$S$. Figure~\ref{fig:lexselschema} shows this process
schematically. %%%\todo{Can we try to bring the figure closer?}
\todomlf{R3 says: ``it would have been nicer to see an example of one sentence going through the pipeline''. Can we add this in the text?}
%Taking our previous example for the parallel corpus, Table~\ref{table:ex-monocorp} 
%presents an example of five possible samples.

@@

suitable normalisation:\todomlf{R2 says ``Nevertheless, I've got problems to understand (4), in particular the paths in post-lexsel step. Some better explanation could be offered maybe by saying that (4) evaluates all possible solutions in the language model.''. Don't know what to do here. }

@@

\todomlf{R1 complains: ``the actual maximum entropy method itself is described very briefly (3 sentences at the end of Section 2)''}
\section{Experimental setting}

@@

Evaluation will be performed using four language pairs. These pairs
have been selected as they include languages with different
morphological complexity, and different amounts of resources available
--- although for all pairs there is a parallel corpus available
for evaluation (see Section~\ref{ss:measures}). 
% Table~\ref{table:lang-pair-stats} 
%presents statistics about the development status of each of the language 
%pairs chosen, including the number of bilingual dictionary entries, coverage of
%test corpora and translational ambiguity of the bilingual dictionary. Translation
%ambiguity is provided for both all words, and for only ambiguous open-category 
%words. In the experiments, lexical-selection rules are learnt for the open categories
%of adjective, noun and verb.

@@

\item[Breton--French:] Development of the Breton--French pair has been
  described in \citep{tyers10b}. The bilingual % tyers09b
  dictionaries were not built with polysemy in mind from the outset,
  but some entries were added later to start work on lexical
  selection. The version used in this paper is SVN revision
  41375.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-br-fr}}
%  This is the only available machine translation system between Breton
%  and French. It has been developed part-time over a number of years.
\todomlf{R1 says ``whole section is rather length especially the revision numbers and URLs of the specific translations systems need not be mentioned in my opinion. The space could rather be used to elaborate in the description of the method in Sec 2.''}

@@

\item[English--Spanish:] The English--Spanish pair was developed from
  a combination of the English--Catalan and Spanish--Catalan pairs, and
%  over a period of around four months. The pair
%has been used in other
%  lexical selection experiments \citep{sanchez07a}\comment{FSM: Sure?
%    It was on en-ca and in that paper no experimental results are
%    shown. This paper would be better mentioned in the introduction
%    (it mentions the use of TL info for lexical selection.},
  contains a number of entries in the bilingual dictionary with more
  than one translation. %The most-frequent translation, as judged by
%  the language pair developer, is marked as the \emph{default}
%  translation and non-default translations were added according to
%  frequency. \todo{Default is explained for some LPs only, why?}  
  The
  version used in this paper is SVN revision
  41387.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-en-es}}\todomlf{4-month mention removed in response to R1}
\end{description}

@@

% see section~\ref{sec:ler}. 
\begin{figure*}
\begin{center}
\setlength\tabcolsep{5.2pt}%
 \begin{tabular}{|l|llllll|}
   \hline
   \multicolumn{7}{|c|}{L'estiu \'{e}s una estaci\'{o} llarga}\\
   \hline
   $S$        & \emph{el} & \emph{estiu} & \emph{ser} & \emph{un} & \emph{estaci\'{o}} & \emph{llarg} \\
   \hline
   $T_s(s_i)$        & \{the\} & \{summer\} & \{be\} & \{a\} & \{station, season\} &  \{long, lengthy\} \\
   \hline 
   $T_r(s_i)$        & \{the\} & \{summer\} & \{be\} & \{a\} &  \{season\} & \{long\} \\
   \hline
   $T_t(s_i)$        & \{the\} & \{summer\} & \{be\} & \{a\} &  \{station\} & \{long\} \\
   \hline
   $\textrm{amb}(s_i)$ & 0     & 0          & 0      & 0     & 1            & 1 \\
   \hline
   $\mathrm{diff}(T_r(s_i), T_t(s_i))$ & 0     & 0          & 0      & 0     & 1            & 0 \\
   \hline
 \end{tabular}
 \caption{An example input sentence in Catalan and the three sets of
   English translations used for calculating the lexical-selection
   error rate. The source sentence $S = (s_1, s_2, \ldots, s_{|S|})$
   has two ambiguous words, \emph{estaci\'{o}} and \emph{llarg} (\(\mathrm{amb}(s_i)=1\), eq.~(\protect\ref{eq:amb})). There is one difference (\(\mathrm{diff}(T_r(s_i),T_t(s_i))=1\), eq.~(\protect\ref{eq:diff}))
   between the reference set $T_r(s_i)$ and the test set $T_t(s_i)$ of
   translations; thus, the error rate for this sentence is
   50\%.}
\label{fig:ler-example}
\end{center}
\end{figure*}

@@

The lexical-selection error rate is the fraction of times the given
system chooses a translation for a word which is not  the one found in an
annotated reference. The process uses a source-language sentence, $S =
(s_1, s_2, \ldots, s_{|S|})$ and three functions. The first function,
$T_s(s_i)$, returns all possible translations of $s_i$ according to
the bilingual dictionary. The second function, $T_t(s_i)$, returns the
translations of $s_i$ selected by the lexical-selection module:
$T_t(s_i) \subseteq T_s(s_i)$; and usually $|T_t(s_i)| = 1$. If the
lexical-selection module returns more than one translation, the first
translation is selected. 
%! \comment{FSM: How many times does this
%!  happens?. I would move this last sentence to an appendix.}  
The
function $T_r(s_i)$ returns the set of reference translations which
are acceptable for $s_i$ in sentence $S$.\footnote{Depending on how
  the reference is built, the set returned by $T_r(s_i)$ % may be
  % incomplete in the sense that it 
  may not include all possible
  acceptable translations.} For a
single sentence, we define the lexical selection error rate (LER) of
that sentence as\todomlf{R3 says: ``The authors use LER, but F-score measures seem more appropriate to estimate the performance of the disambiguation module. Moreover, the constraint of the disambiguation module to generate only one word when evaluating seems extreme.'' He don't get it.}
\begin{equation}
\mathrm{LER} = \frac{\sum_{i=1}^{|S|} \mathrm{amb}(s_i)\; \mathrm{diff}(T_r(s_i), T_t(s_i))}{\sum_{i=1}^{|S|} \mathrm{amb}(s_i)},
\label{eq:ler}
\end{equation}
where
\begin{equation}
\mathrm{amb}(s_i) = \left \{ \begin{matrix}
                 1 & \mbox{if } |T_s(s_i)| > 1   \\
                 0 & \mbox{otherwise} \end{matrix} \right.
\label{eq:amb}
\end{equation}
 tests if a word is ambiguous, and the  function 
\begin{equation}
\mathrm{diff}(T_r(s_i), T_t(s_i)) = \left \{ \begin{matrix}
                 1 & \mbox{if } T_r(s_i) \cap T_t(s_i) = \emptyset   \\
                 0 & \mbox{otherwise} \end{matrix} \right.
\label{eq:diff}
\end{equation}
states that there is a difference if the intersection between the set
of reference translations $T_r(s_i)$ and the set of translations from
the lexical selection module $T_t(s_i)$ is empty.  Recall that,\todomlf{R1 says ``this has already been mentioned before''. Don't know what to do}
although $T_t(s_i)$ returns a set, this set will be a singleton, as
when the lexical-selection module returns more than one translation
the first one will be selected.\footnote{In practice this does not happen
as each ambiguous word has a \emph{default} translation.}\todomlf{R1 complains: ``why mention if it doesn't happen in practice? Relevance?'' What should we do?}

@@

The table in Figure~\ref{fig:ler-example} gives an
overview of the inputs.\todomlf{Caption of figure extended in response to R3} In the description it is assumed that the
reference translation has been annotated by hand. However, hand
annotation is a time-consuming process, and was not possible. A
description of how the reference was built is given in
Section~\ref{sec:eval-corpus}.

\subsubsection{Machine translation performance} 
%! \comment{This part of the text needs compressing and merging} 
This is an extrinsic evaluation, which ideally would test how much the
system improves as regards an approximate measurement of final
translation quality in a real system. For this task, we use the
widely-used {\sc bleu} metric \citep{papineni02}.\todomlf{BLEU reference advanced, and wrong section link removed} This is not ideal for evaluating the
task of a lexical selection module as the performance of the module
will depend greatly on (a) the coverage of the bilingual dictionaries
of the RBMT system in question, and (b) the number of reference
translations. It is also worth noting that successful lexical
selections may not lead to successful translations due to inadequate
transfer of morphological features.
% For example, if
% we translate the Spanish phrase \emph{La gente le dice que no venga}
% `The people tell him not to come', where \emph{decir} can be
% translated as `say' or `tell' in English, and our machine translation
% system generates `The people \emph{tells} him not to come', it would
% be counted the same as if it were `The people \emph{says} him not to
% come', although the lexical selection made was more
% adequate. Generation errors --- where a word is found in the bilingual
% dictionary, but not in the target language morphological dictionary
% --- may also lead to the same problem.  
The {\sc bleu} metric  is
included only as it is commonly used to evaluate MT systems.

@@

  After this process, we selected 1,000 sentence pairs at random for
  testing (\texttt{test}), 1,000 for development
  (\texttt{dev})\footnote{The development corpus was used for checking
    the value for frequency pruning of features.} and left the
  remainder for training.
%This is a smaller
%number of sentences than used in typical evaluations (such as the WMT
%series, \cite{callisonburch2012}), but was motivated by the fact that
%the smallest corpus (for Breton--French) after sentence discarding was
%only little over 4,000 sentences long. 
  Table~\ref{table:input-corp} gives statistics about the size of the
  input corpora, and how many sentences were left after processing for
  testing, training and development. %\todo{Why do ambiguity numbers
%    change so much from the paper?} 
Table~\ref{table:test-corp} gives
  information about the test corpora. 
%! \todo{FSM: One may wonder which
%!    would be the results when the TL side of the training corpus is
%!    used to train the models in the usual way.}

@@

As we are working with binary features, we use the implementation of
generalised iterative scaling available in the {\sc
  yasmet}\footnote{\url{http://www-i6.informatik.rwth-aachen.de/web/Software/YASMET.html};
  the compilable version we used is available as part of the Apertium
  \texttt{lex-tools} package,
  \url{http://downloads.sourceforge.net/project/apertium/apertium-lex-tools/apertium-lex-tools-0.1.0.tar.gz}.}
to calculate the feature weights.  
\todomlf{R3 says: ``The proposed maximum entropy or log-linear model is trained in a surprising way. In the early days of maximum entropy models, those were trained using the GIS algorithm. However, it has been proved in the literature that GIS has clear disadvantages with respect to other batch gradient descent algorithms. For instance, it does not allow for the use of regularization [...], and it is extremely slow to converge, which yields worse models than other methods. The smoothing/regularization of the model is also poor since it is simply feature pruning instead of regularization.''}
After learning the feature sets and
weights, we compute the evaluation measures described in
Section~\ref{ss:measures}. There is an option to remove events 
$(s, t, c)$ which occur
less than a certain number of times in the training corpus. This is
referred to as the feature pruning frequency threshold --- features
occuring less than the threshold are discarded. The value was set
experimentally. Values of between two and seven were tested, and the
ones which provided the best improvement on the development corpus
were selected; they happen to come close to the rule-of-thumb value of
five that \citet[p.~596]{Manning99b} found to be effective.
Table~\ref{table:maxent-features} shows the number of features that
have eventually been used for each language pair.

@@

\paragraph{Acknowledgements:} We acknowledge support from the Spanish Ministry of Industry and Competitiveness through project Ayutra (TIC2012-32615) and from the European Commission through project Abu-Matran (FP7-PEOPLE-2012-IAPP, ref.\ 324414) and thank all three anonymous referees for very useful comments on the paper.\todomlf{Fran: add your acknowledgements}
